\documentclass[11pt,a4paper]{article}

\usepackage{xcolor}
\usepackage[colorlinks=true, linkcolor=black!50!blue, urlcolor=blue, citecolor=blue, anchorcolor=blue]{hyperref}
\usepackage[font=small,labelfont=bf,margin=0mm,labelsep=period,tableposition=top]{caption}
\usepackage[a4paper,top=3cm,bottom=2.5cm,left=2.5cm,right=2.5cm,bindingoffset=0mm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{dsfont}
\usepackage{pifont}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{makeidx}
\usepackage{simplewick}
\usepackage{placeins}
\usepackage{bbold}
\usepackage{braket}
\graphicspath{{./figs/}}

\input{commands}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\ldd}[1]{\textcolor{red}{\textbf{Luigi: #1}}}
\newcommand{\ac}[1]{\textcolor{red}{\textbf{Amedeo: #1}}}

\begin{document}
\newgeometry{top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm,bindingoffset=0mm}
\vspace{-2.0cm}
\begin{flushright}
Edinburgh 2025/x
\end{flushright}
\vspace{0.3cm}

\begin{center}
  {\Large \bf On the Origin of PDF Uncertainties}
%  {\Large \bf Diagnostic Tools for Parton Distribution Function Fits}
%  {\Large \bf Towards a Quantitative Understanding of PDF Fits and their Uncertainties}
  
  \vspace{1.0cm}

\vspace{1.1cm}

  Amedeo Chiefa, Luigi Del Debbio and Richard Kenway

  \vspace{0.2cm}

  {\it \small
    The Higgs Centre for Theoretical Physics, \\ 
    School of Physics and Astronomy, The University of Edinburgh,\\
    Peter Guthrie Tait Road, Edinburgh EH9 3FD, United Kingdom\\[0.1cm]
  }
  \vspace{0.7cm}
\end{center}

\begin{abstract}
  Parton Distribution Functions (PDFs) play a central role in describing
  experimental data at colliders and provide insight into the 
  structure of nucleons. As the LHC enters an era of high-precision measurements, a robust
  PDF determination with a reliable uncertainty quantification has become
  mandatory in order to match the experimental precision. The NNPDF
  collaboration has pioneered the use of Machine Learning (ML) techniques for PDF
  determinations, using Neural Networks (NNs) to parametrise the unknown PDFs 
  in a flexible and unbiased way. The NNs are then trained on experimental data
  by means of stochastic gradient descent algorithms. The statistical robustness 
  of the results is validated by extensive closure tests using synthetic data.
  In this work, we develop a theoretical framework based on the
  Neural Tangent Kernel (NTK) to analyse the training dynamics of neural
  networks. This approach allows us to derive, under precise assumptions, an
  analytical description of the neural networks evolution during training,
  enabling a quantitative understanding of the training process. 
  Having an analytical handle on the training dynamics allows us to clarify the
  role of the NN architecture and the impact of the experimental data in a 
  transparent way. Similarly, we are able to describe the evolution of the covariance of
  the NN output during training, providing a quantitative description of how
  uncertainties are propagated from the data to the fitted function.
  Interestingly, the methodology developed in this work can be used to understand the 
  minimization of a loss function for any kind of parametrization, thereby 
  providing a unified framework to compare different PDF determinations, like,
  \eg, fits based on a particular functional form. While our results are {\em not}\ 
  a substitute for PDF fitting, they do provide a powerful diagnostic tool to
  assess the robustness of current fitting methodologies.
  Beyond its relevance for particle physics phenomenology, our analysis of 
  PDF determinations provides a testbed to apply theoretical ideas about the 
  learning process developed in the ML community. As seen in applications from 
  other domains, we find that our results deviate from the simple picture 
  of the \textit{lazy training} regime discussed in the ML literature.
\end{abstract}

\newpage

\tableofcontents
\clearpage

% Main Sections
\input{introduction.tex}
\input{initialisation.tex}
\input{ntk_training.tex}
\input{lazy_training.tex}
\input{conclusions.tex}

\section*{Acknowledgements}
LDD is suppoorted by an STFC Consolidated Grant (ST/T000600/1, ST/X000494/1). 

\newpage

% Appendices
\appendix
\input{dataset.tex}
\input{ntk_arch_dep.tex}
\input{cutoff.tex}

% Bibliography
\bibliographystyle{unsrt}
\bibliography{ntk.bib}

\end{document}